---
title: "W2_data_prep_lab"
author: "Miao Li"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
<span style="color:blue"><font size="4"> FYI, a friend once said:</font> </span>

> This task is <span style="color:red">__NOT__</span> as difficult 
> as figuring out why $E=mc^2$.

## I. The following codes load the necessary packages for this task 

```{r load package, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(readxl)
library(tidyverse)
library(labelled)
library(DT) 
```
NOTE: You will first install these packages before loading them using the `library()` function. You can use the following functions to install these packages:
 `install.packages("readxl")`
 `install.packages("labelled")`
 `install.packages("data.table")`
 `install.packages("tidyverse")`

## II. Read in Data

* You need to specify your own directory leading to your data file
* Note that in the directory the slashes are forward, __NOT__ backward!
* You can also take a quick peek at the first few rows of the data using the function `head()`

```{r echo=TRUE}
mysurvey <- read_excel("C:/Users/miaol/OneDrive - Clemson University/My Teaching/Clemson/SSCI_8010/Example_data/Survey_results.xlsx")    
head(mysurvey)    

```

## III. Rename the variables and give them labels

* The current variable names are too long (they are the original survey questions asked)! 
```{r}
names(mysurvey)  
# this function allows you to take a look at the current variable names. 
# Indeed, they are LOOOOOOOOOOOONG!
```

* rename them with shorter names (Rule of Thumb: 8 letters or less)
* and create an ID variable with the name "ID"
* then peek your data again. see what has changed.
```{r}
names(mysurvey) <- c("belief", "gender", "age", "reliattend", "belief12","alcohol","presex","homosex","polaffi")
mysurvey <- rowid_to_column(mysurvey, "ID")
head(mysurvey)
```
* Shorter names mean convenience, but with loss of information.
* Create a label for each variable to document what the variable is about
```{r}
var_label(mysurvey) <- list(
  ID         = "Individual ID",
  belief     = "Are you a Christian?",
  gender     = "What is your sex?",
  age        = "How old are you?",
  reliattend = "How often do you go to a church?",
  belief12   = "Were you a Christian when you were 12 years old?",
  alcohol    = "Largest number of drinks in one occasion in the past 30 days",
  presex     = "How much agree with ‘premarital sex is always wrong'?",
  homosex    = "How much agree with same-sex marriage?",
  polaffi    = "Political affiliation: Rep, Demo, or Ind?"
)
```

* if you want to know what change has been done, you can go back to RStudio and type in the function ```View()``` to get a tabular view of the data set (i.e., like what you would see in Excel)
* you will see that under each variable name there is a label with detailed info about what this variable is about
```{r}
datatable(mysurvey)
```

## IV. Clean the Data
The following chunk of code does three things:  \
1. Re-order the levels for the ordinal-level variable "`reliattend`", which is stored in R as a [factor variable](https://towardsdatascience.com/introduction-to-factors-in-r-dd752f732c94)  \
2. Re-code the value "`Not Sure`" for the variable "`belief`" in to missing values \
3. Generate a numeric version of the ordinal variable "`reliattend`", name it as "`attend`" \
__NOTE:__ The code completed all three tasks in one pipeline. The pipe operator sign `%>%` represents the link of the pipe. It tells R that the command is not finished yet and the next step is based on the previous step. Read more [here](https://style.tidyverse.org/pipes.html).
```{r}
mysurvey <-  mysurvey %>%
  mutate_if(is.character, as.factor) %>% 
  mutate(reliattend = factor(reliattend, levels = c("Rarely or never", "Less than monthly", "Monthly but not weekly", "Weekly or more"))) %>% 
  mutate(belief = fct_recode(belief, NULL = "Not Sure")) %>% 
  mutate(attend = as.numeric(reliattend)) 
var_label(mysurvey$attend) <- "Religious attendance (numeric)"
```

* label values of the new variable `attend` as well as the numeric variables `presex` and `homosex`
```{r}
val_labels(mysurvey$attend) <- c("Rarely or never"   =1, 
                                "less than monthly"  =2, 
                                "monthly not weekly" =3, 
                                "weekly or more"     =4) 
val_labels(mysurvey[,c("presex","homosex")]) <- c("strongly disagree" =1,
                                                  disagree            =2,
                                                  neutral             =3,
                                                  agree               =4,
                                                  "strongly agree"    =5)
```

* then check the labels

```{r, echo=TRUE}
options(width = 300) #this is to set the page width wider
val_labels(mysurvey[,c("attend", "presex","homosex")])
```

* lastly, create a new variable for binge drinking 
* (definition of binge drinking: >=4 drinks/occasion for female, >=5 drinks/occasion for male)
* label the variable with appropriate info

```{r}
mysurvey <- mysurvey %>% 
  mutate(binge = case_when(
    alcohol>=4 & gender=="Female" & !is.na(alcohol) ~ 1,
    alcohol>=5 & gender=="Male"   & !is.na(alcohol) ~ 1,
    alcohol<4  & gender=="Female"                   ~ 0,
    alcohol<5  & gender=="Male"                     ~ 0)
    )
var_label(mysurvey$binge) <- "Binge drinking"
val_labels(mysurvey$binge) <- c(yes=1, no=0)
head(mysurvey$binge)
```

## V. Construct a data dictionary 
* A data dictionary includes the meta-data of the data set
* Save the dictionary into a different data frame and name it as `dat_dic`

```{r}
dat_dic <- lookfor(mysurvey)  
datatable(dat_dic)
```
